### 2024-08-19 paper.tex
move linalg section
fixup linalg section

@@

\section{Simplifying the definition of a Springer fiber at a Slodowy slice}
\par Let \(\mathcal{N}_m \subseteq \gl_m(\mathbb{C})\) be the subset consisting of nilpotent elements.
For \(X \in \mathcal{N}_m\) and \(a, b \in \mathbb{C}^m\), let
\[A_{X, a, b} = \begin{pNiceArray}{ccc|ccccc}
 & & & & & & & \vert \\
 & X & & & & & & b    \\
 & & & & & & & \vert \\
\hline
\text{---} & a & \text{---} & 0 \\
& & & 1 & 0 \\
& & &  & 1 & \ddots \\
& & &  &   & \ddots & 0 \\
& & &  &   &  & 1 & 0
\end{pNiceArray} \in \mathfrak{gl}_{m + n}(\mathbb{C}).\]
By the definition given in the previous section, we have 
\[\widetilde{\mathcal{N}}_{m, n} = \{(\mathfrak{b}, X) : X \in \mathfrak{b}\} \subseteq \mathcal{B}_{m + n} \times \mathcal{S}_{m,n}',\]
and the Springer fiber at the \(n\)-Slodowy slice at a nilpotent \(X \in \gl_m\) is
\[\pi_{m,n}^{-1}(X) = \{(\mathfrak{b}, A_{X, a, b}) : A_{X, a, b} \in \mathfrak{b}, A_{X, a, b} \in \mathcal{S}_{m, n}'\} \] 
\[= \{(\mathfrak{b}, A_{X, a, b}) : A_{X, a, b} \in \mathfrak{b}, A_{X, a, b} \textrm{nilpotent}\}\]
\[\cong \{(\mathfrak{b}, a, b) : A_{X, a, b} \in \mathfrak{b}, A_{X, a, b} \textrm{ nilpotent}\}.\]
We will make one last simplification to this by using a correspondence between Borel subalgebras and complete flags.
TODO: fix the next few paragraphs, they are out of context.
\par We have \(\mathcal{M} = \{AHA^{-1} : A \in \GL_{m + n}(\mathbb{C})\}\), where \(H \subseteq \gl_{m + n}(\mathbb{C})\) is the set of upper triangular matrices.
We say a map \(X : \gl_{m + n} \to \gl_{m + n}\) preserves a flag \(V_0 \subseteq \cdots \subseteq V_{m + n}\) if \(XV_i \subseteq V_i\) for each \(i\).
Let \(E_0 \subseteq \cdots \subseteq E_{m + n}\) be the standard flag of \(\mathbb{C}^{m + n}\).
Since \(H\) is the set of \(X\) which preserve \(E\),
\[\mathcal{M} = \{\{X : \forall i. \; X(AE_i) \subseteq AE_i\} : A \in \GL_n(\mathbb{C})\}.\]
So, for \(X \in \mathcal{N}\),
\[\pi^{-1}(X) \cong \{(V, a, b) : \forall i. \; A_{X, a, b} V_i \subseteq V_{i - 1}\}.\]

@@

\section{Self-contained linear algebra facts}
\subsection{The centralizer of a nilpotent matrix}
\begin{definition}
    A matrix \(Y\) is \emph{Toeplitz} if it is constant along bands parallel to the main diagonal.
    That is, \(\forall i,j,k. \; Y_{ij} = Y_{i + k, j + k}\).
\end{definition}
\begin{definition}
    An \(m \times n\) matrix \(Y\) is \emph{lower-left Toeplitz} if it is Toeplitz and, in addition, only the lower 
\end{definition}
For nilpotent \(X\) in Jordan form and \(a\) in `normalized' form (i.e., with at most one nonzero element, which is a one), we will find the centralizer of \(A_{X, a, b}\) in 
\[\{(A, I) : A \in \gl_m\} \subseteq \gl_{m + n}.\]
Note that an element of the form \((A, I)\) commutes with \(A_{X, a, b}\) if and only if \(A\) commutes with \(X\), and \(aA = a\), and \(Ab = b\).
So, we just have to find 
\[\{A \in \gl_m : AX = XA, aA = a, Ab = b\}.\]
We begin by finding 
\[\{A \in \gl_m : AX = XA\}.\]
We write the shape of \(X\) as \(\lambda = (\lambda_1, ..., \lambda_k)\), so that 
\[X = \begin{pmatrix}
    J_{\lambda_1} & \\
    & \ddots \\
    & & J_{\lambda_k}
\end{pmatrix}.\]
Then we write \(A\) as the block matrix
\[A =: \begin{pmatrix}
    A_{11} & \cdots & A_{1k}\\
    \vdots & & \vdots\\
    A_{k1} & \cdots & A_{kk}
\end{pmatrix},\]
where \(A_{ij}\) is a block of size \(\lambda_i \times \lambda_j\).
\[XA = \begin{pmatrix}
    J_{\lambda_1}A_{11} & \cdots & J_{\lambda_1}A_{1k}\\
    \vdots & & \vdots\\
    J_{\lambda_k}A_{k1} & \cdots & J_{\lambda_k}A_{kk}
\end{pmatrix}, \textrm{ and } AX = \begin{pmatrix}
    A_{11} J_{\lambda_1} & \cdots & A_{1k} J_{\lambda_k}\\
    \vdots & & \vdots\\
    A_{k1} J_{\lambda_1} & \cdots & A_{kk} J_{\lambda_k}
\end{pmatrix}.\]
So, the constraint that \(XA = AX\) is simply saying that 
\[\forall i,j. \; J_{\lambda_i} A_{ij} = A_{ij} J_{\lambda_j}.\]
It's easy to see that left multiplication by \(J_{\lambda_i}\) is just a down-shift by one, and right multiplication by \(J_{\lambda_j}\) is a left-shift by one.
This means that \(A_{ij}\) is constrained to be a ``lower-left-Toeplitz'' matrix.
A Toeplitz matrix is one which is constant along diagonal bands: \(\forall ijk. \; x_{ij} = x_{i + k,j + k}\).
A lower-left-Toeplitz matrix is one which is all zeroes except for the bottom-left corner; that is, the bands start in the bottom-left corner, and continue until hitting the band which includes the lower-right corner or the band which includes the upper-left corner (whichever comes first).
In other words, an \(n \times m\) lower-left-Toeplitz matrix is one which satisfies \(x_{ij} = 0\) for all \(i,j\) with \(i < j\) and all \(i,j\) with \(i - j < m - n\).
In yet other words, a matrix is lower-left-Toeplitz if it is Toeplitz, satisfies \(x_{1j} = 0\) for \(j \neq 1\), and satisfies \(x_{in} = 0\) for \(i \neq m\).
Anyway, it is easy to see that the lower-left-Toeplitz matrices are exactly those matrices \(A\) such that left-shifting \(A\) by one has the same result as down-shifting by one.
\par So, we have found the set 
\[C_1 := \{A \in \gl_m : AX = XA\}.\]
It is just the set of \(A = [A_{ij}]_{ij}\), where each \(A_{ij}\) is lower-left-Toeplitz.
\par Let \(v_{ij}\) be the leftmost column of \(A_{ij}\), so that
\[A_{ij} = \begin{pmatrix}
    v_{ij} \downshift 0 & \cdots & v_{ij} \downshift [\lambda_j - 1]
\end{pmatrix}.\]
Since \(A_{ij}\) is lower-left-Toeplitz, \(v_{ij}\) is of the form 
\[v_{ij} = \begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i, \lambda_j)}\) can be chosen freely.
Now we determine which matrices of this form satisfy \(aA = a\).
For simplicity, we will instead find the matrices \(A\) such that \(aA = 0\) (so that \(a(A + I) = a\)).
(Note that \(I \in C_1\), and \(C_1\) is closed under addition, so this is really the same thing.)
\par In the case that \(a = 0\), clearly every \(A\) works.
Otherwise, let \(i_0, j_0\) be such that \(a_{i_0,j_0} = 1\).
Now, clearly the constraint that \(aA = 0\) is just saying that the \((i_0,j_0)\)th row of \(A\) must be zero.
That is, for each \(j\) the \(j_0\)th row of \(A_{i_0j}\) must be zero.
This just requires that for each \(j\), we must have 
\[v_{ij} = \begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i - j_0, \lambda_j)}\) can be chosen freely.
So, we have now found the set 
\[C_2 := \{A \in \gl_m : AX = XA, aA = a\}.\]
It is the matrices of the form \(I + A\), where \(A_{ij}\) is a block matrix of size \(\lambda_i \times \lambda_j\) with 
\[A_{ij} = T(v_{ij}) = T\begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i, \lambda_j)}\) in the case \(i \neq i_0\) and \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i - j_0, \lambda_j)}\) if \(i = i_0\).

>>>>

\section{Some linear algebra facts}
In this section we prove some facts about linear algebra that will be useful to us later.
\subsection{The centralizer of a nilpotent matrix}
\begin{definition}
    A matrix \(Y\) is \emph{Toeplitz} if it is constant along bands parallel to the main diagonal.
    That is, \(\forall i,j,k. \; Y_{ij} = Y_{i + k, j + k}\).
\end{definition}
\begin{definition}
    An \(m \times n\) matrix \(Y\) is \emph{lower-left Toeplitz} if it is Toeplitz and, in addition, we have \(y_{n - i, j - 1} = 0\) whenever \(i + j \geq \min(m,n)\).
\end{definition}
That is, \(Y\) is lower-left Toeplitz if it is Toeplitz, and the only nonzero entries are those with Manhattan distance less than \(\min(m,n)\) from the lower-left corner.

\begin{lemma}
    Let \(\lambda = (\lambda_1, ..., \lambda_k)\) be a partition.
    The centralizer of \(J(\lambda)\) is the set of matrices 
    \[M = \begin{pmatrix}
        M_{11} & \cdots & M_{1k}\\
        \vdots &        & \vdots\\
        M_{k1} & \cdots & M_{kk}
    \end{pmatrix},\]
    where each \(M_{ij}\) is a \(\lambda_i \times \lambda_j\) matrix, such that each \(M_{ij}\) is lower-left Toeplitz.
\end{lemma}
\begin{proof}
    Let 
    \[M = \begin{pmatrix}
        M_{11} & \cdots & M_{1k}\\
        \vdots &        & \vdots\\
        M_{k1} & \cdots & M_{kk}
    \end{pmatrix}.\]
    We need to show that \(J(\lambda)M = MJ(\lambda)\) if and only if each \(M_{ij}\) is lower-left Toeplitz.
    \par We have 
    \[J(\lambda)M = \begin{pmatrix}
        J_{\lambda_1}M_{11} & \cdots & J_{\lambda_1}M_{1k}\\
        \vdots & & \vdots\\
        J_{\lambda_k}M_{k1} & \cdots & J_{\lambda_k}M_{kk}
    \end{pmatrix}, \textrm{ and } MJ(\lambda) = \begin{pmatrix}
        M_{11} J_{\lambda_1} & \cdots & M_{1k} J_{\lambda_k}\\
        \vdots & & \vdots\\
        M_{k1} J_{\lambda_1} & \cdots & M_{kk} J_{\lambda_k}
    \end{pmatrix}.\]
    So, we have \(J(\lambda)M = MJ(\lambda)\) if and only if \(\forall i,j. \; J_{\lambda_i} M_{ij} = M_{ij} J_{\lambda_j}\).
    Multiplying on the left by \(J_{\lambda_i}\) just shifts each row down by one, and multiplying on the right by \(J_{\lambda_j}\) shifts each column left by one.
    The matrices for which left-shifting gives the same result as down-shifting are exactly the lower-left Toeplitz matrices.
\end{proof}
For nilpotent \(X\) in Jordan form and \(a\) in `normalized' form (i.e., with at most one nonzero element, which is a one), we will find the centralizer of \(A_{X, a, b}\) in 
\[\{(A, I) : A \in \gl_m\} \subseteq \gl_{m + n}.\]
Note that an element of the form \((A, I)\) commutes with \(A_{X, a, b}\) if and only if \(A\) commutes with \(X\), and \(aA = a\), and \(Ab = b\).
So, we just have to find 
\[\{A \in \gl_m : AX = XA, aA = a, Ab = b\}.\]

It is just the set of \(A = [A_{ij}]_{ij}\), where each \(A_{ij}\) is lower-left-Toeplitz.
\par Let \(v_{ij}\) be the leftmost column of \(A_{ij}\), so that
\[A_{ij} = \begin{pmatrix}
    v_{ij} \downshift 0 & \cdots & v_{ij} \downshift [\lambda_j - 1]
\end{pmatrix}.\]
Since \(A_{ij}\) is lower-left-Toeplitz, \(v_{ij}\) is of the form 
\[v_{ij} = \begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i, \lambda_j)}\) can be chosen freely.
Now we determine which matrices of this form satisfy \(aA = a\).
For simplicity, we will instead find the matrices \(A\) such that \(aA = 0\) (so that \(a(A + I) = a\)).
(Note that \(I \in C_1\), and \(C_1\) is closed under addition, so this is really the same thing.)
\par In the case that \(a = 0\), clearly every \(A\) works.
Otherwise, let \(i_0, j_0\) be such that \(a_{i_0,j_0} = 1\).
Now, clearly the constraint that \(aA = 0\) is just saying that the \((i_0,j_0)\)th row of \(A\) must be zero.
That is, for each \(j\) the \(j_0\)th row of \(A_{i_0j}\) must be zero.
This just requires that for each \(j\), we must have 
\[v_{ij} = \begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i - j_0, \lambda_j)}\) can be chosen freely.
So, we have now found the set 
\[C_2 := \{A \in \gl_m : AX = XA, aA = a\}.\]
It is the matrices of the form \(I + A\), where \(A_{ij}\) is a block matrix of size \(\lambda_i \times \lambda_j\) with 
\[A_{ij} = T(v_{ij}) = T\begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i, \lambda_j)}\) in the case \(i \neq i_0\) and \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i - j_0, \lambda_j)}\) if \(i = i_0\).

@@

\section{Simplifying the definition of a Springer fiber at a Slodowy slice}
\par Let \(\mathcal{N}_m \subseteq \gl_m(\mathbb{C})\) be the subset consisting of nilpotent elements.
For \(X \in \mathcal{N}_m\) and \(a, b \in \mathbb{C}^m\), let
\[A_{X, a, b} = \begin{pNiceArray}{ccc|ccccc}
 & & & & & & & \vert \\
 & X & & & & & & b    \\
 & & & & & & & \vert \\
\hline
\text{---} & a & \text{---} & 0 \\
& & & 1 & 0 \\
& & &  & 1 & \ddots \\
& & &  &   & \ddots & 0 \\
& & &  &   &  & 1 & 0
\end{pNiceArray} \in \mathfrak{gl}_{m + n}(\mathbb{C}).\]
By the definition given in the previous section, we have 
\[\widetilde{\mathcal{N}}_{m, n} = \{(\mathfrak{b}, X) : X \in \mathfrak{b}\} \subseteq \mathcal{B}_{m + n} \times \mathcal{S}_{m,n}',\]
and the Springer fiber at the \(n\)-Slodowy slice at a nilpotent \(X \in \gl_m\) is
\[\pi_{m,n}^{-1}(X) = \{(\mathfrak{b}, A_{X, a, b}) : A_{X, a, b} \in \mathfrak{b}, A_{X, a, b} \in \mathcal{S}_{m, n}'\} \] 
\[= \{(\mathfrak{b}, A_{X, a, b}) : A_{X, a, b} \in \mathfrak{b}, A_{X, a, b} \textrm{nilpotent}\}\]
\[\cong \{(\mathfrak{b}, a, b) : A_{X, a, b} \in \mathfrak{b}, A_{X, a, b} \textrm{ nilpotent}\}.\]
\par Using our correspondence between \(\mathcal{B}_{m + n}\) and complete flags of \(\mathbb{C}^{m + n}\), we obtain 
\[\pi_{m,n}^{-1}(X) \cong \{(V, a, b) : \forall i. \; A_{X, a, b} V_i \subseteq V_{i - 1}\}.\]

@@
